---
title: "Reproducible Research"
author: "Hugh Li"
date: "June 12, 2015"
output: html_document
---
## Loading and preprocessing the data

```{r,echo=TRUE}
if(!file.exists("./data")){dir.create("./data")}
url<-'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'
download.file(url,destfile = './a.zip',method='curl')
unzip('a.zip')
s<-read.csv("activity.csv")  # read data
s$date<-as.Date(s$date, "%Y-%m-%d") #transform class to date
```

## What is mean total number of steps taken per day?
```{r, echo=TRUE}
m<-with(s,tapply(steps,list(date),sum,na.rm=TRUE)) #calculate daily steps
hist(m,xlab = 'The total number of steps per day',main='Histogram of total steps per day')
mean(m,na.rm = TRUE)    
median(m,na.rm=TRUE)
```

## What is the average daily activity pattern?
```{r}
n<-with(s,tapply(steps,list(interval),mean,na.rm=TRUE))
plot(x=names(n),n,xlab='Interval Identifier',ylab='Average of steps across all days',type='l',main="Average daily activity pattern")

# interval which contains the maximum number of steps
names(n)[match(max(n),n)]
```

##Imputing missing values 
```{r}
# count missing values
sum(is.na(s$steps)) 

# replace missing values with the mean for that day.
x<-with(s,tapply(steps,list(date),mean,na.rm=TRUE)) #calculate average steps in a specific day 
x[is.na(x)]<-0 # NaN values in x, coz missing values for a whole day. 
x1<-rep(x,each=288) # 17568 rows with values of average steps on corresponding days
s.new<-s
for (i in c(1:17568)){
  if(is.na(s.new$steps[i])==TRUE){
   s.new$steps[i]=x1[[i]] 
  }
}
m.new<-with(s.new,tapply(steps,list(date),sum,na.rm=TRUE)) #calculate daily steps
hist(m.new,xlab = 'The total number of steps per day',main='New histogram of total steps per day')
mean(m.new,na.rm = TRUE)    
median(m.new,na.rm=TRUE)
```

Imputation methods of replacing missing values with mean of that day don't affect the estimates, which mathematically makes sense.
Imputation increases total daily number of steps. You just add 'new' samples to the original dataset.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lattice)
weekday<-c('Monday','Tuesday','Wednesday','Thursday','Friday')
s.new$w<-factor((weekdays(s.new$date) %in% weekday)+1L,levels=1:2,labels=c("weekend","weekday")) # create a new column indicating weekday or weekend

#n.new<-with(s.new,tapply(steps,list(interval,w),mean,na.rm=TRUE))
s.weekday<-subset(s.new,w=='weekday')
s.weekend<-subset(s.new,w=='weekend')
n.weekday<-with(s.weekday,tapply(steps,list(interval),mean,na.rm=TRUE))
xx<-names(n.weekday)
xx1<-as.numeric(xx)
n.weekday1<-data.frame(xx1,n.weekday,rep('weekday',288))
colnames(n.weekday1)<-c('Interval','ave','week')
n.weekend<-with(s.weekend,tapply(steps,list(interval),mean,na.rm=TRUE))
n.weekend1<-data.frame(xx1,n.weekend,rep('weekend',288))
colnames(n.weekend1)<-c('Interval','ave','week')
f<-rbind(n.weekday1,n.weekend1)
xyplot(f$ave~f$Interval|f$week,xlab="Interval Identifier",ylab="Average of steps",type='l',layout=c(1,2))

```

